---
title: "A Friendly Introduction to Github for Project Version Control"
author: "Hao Ye"
institute: "Health Science Center Libraries, University of Florida"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#09274B",
  base_font_size = "32px", 
  header_h1_font_size = "2.0rem",
  header_h2_font_size = "1.5rem",
  header_h3_font_size = "1.25rem", 
  header_background_content_padding_top = "2rem", 
  header_font_google = google_font("Lato"),
  text_font_google   = google_font("Montserrat", "300", "300i", "500"),
  code_font_google   = google_font("Source Code Pro", "300", "500"), 
  extra_css = list("p" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "ul" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "li" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "pre" = list("margin-top" = "0.5em", 
                              "margin-bottom" = "0.5em")), 
  link_color = "#339944"
)
```
class: inverse, center, middle

# `r rmarkdown::metadata$title`
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$institute`
### (updated: `r Sys.Date()`)

---

# Motivations

Good record-keeping of your projects makes it easier to:
* revert back to previous versions
* interact with collaborators working on the same files
* share your work and have it cited

---

# Learning Outcomes

By the end of the workshop, participants will be able to:
* explain the benefits of version control
* articulate the differences between `git` and `Github`
* create and update repositories using Github Desktop and the Github website
* write descriptive README files
* select and apply a license to a GitHub project

---

# Version Control

Tools/systems to manage changes to files.

--
## So what?

* MS Word has "track changes"
* Dropbox / Google Drive can revert versions

---

# For *project* management

What about?

* changes that are not in MS Word
* restoring a version of the project involving many different files
* information about *who* and *why*
* multiple collaborators working on the same files

---

# Example scenarios

* You change a histogram to a bar plot, and then you are asked to change it back
* Your collaborator makes edits while at a conference, and you need to merge the changes with your own
* You want to experiment with a different dataset / analysis / narrative without duplicating all the files and keeping track of multiple copies of the project

---

# Git




---

# Thanks

* Let me know what content you'd like to see
* Contact me for additional questions or consultation requests!
* Check back in on the libguide for more modules and contact info:
  - https://guides.uflib.ufl.edu/reproducibility
