---
title: "A Friendly Introduction to Github for Project Version Control"
author: "Hao Ye"
institute: "Health Science Center Libraries, University of Florida"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#09274B",
  base_font_size = "32px", 
  header_h1_font_size = "2.0rem",
  header_h2_font_size = "1.5rem",
  header_h3_font_size = "1.25rem", 
  header_background_content_padding_top = "2rem", 
  header_font_google = google_font("Lato"),
  text_font_google   = google_font("Montserrat", "300", "300i", "500"),
  code_font_google   = google_font("Source Code Pro", "300", "500"), 
  extra_css = list("p" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "ul" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "li" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "pre" = list("margin-top" = "0.5em", 
                              "margin-bottom" = "0.5em")), 
  link_color = "#339944"
)
```
class: inverse, center, middle

# `r rmarkdown::metadata$title`
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$institute`
### (updated: `r Sys.Date()`)

---

# Motivations

Good practices in project management make tasks easier:
* revert back to previous versions
* interact with collaborators working on the same files
* share your work and have it cited

---

# Learning Outcomes

By the end of the workshop, participants will be able to:
* explain the benefits of version control
* articulate the differences between `git` and `Github`
* create and update repositories using Github Desktop and the Github website
* write descriptive README files
* select and apply a license to a GitHub project

---

class: inverse, center, middle

# Version Control Concepts

---

# Version Control

Tools/systems to manage changes to files.

--
## So what?

* MS Word has "track changes"
* Dropbox / Google Drive can revert versions

---

# What about *projects*?

* changes that are not in MS Word
* restoring multiple files that depend on each other
* knowing *who* made changes and  *why*
* multiple collaborators working on the same files

---

# Example scenarios

* You change a histogram to a bar plot, and then you are asked to change it back
* Your collaborator makes edits while traveling, and you need to merge their changes with your own changes
* You want to experiment with a different dataset / analysis / narrative without duplicating everything

---

class: inverse, center, middle

# Git and GitHub

---

# Git

* software and version control system
* notoriously unfriendly
  - unintuitive command names
  - obscure error messages

--

**This workshop is NOT about git!**
* check out the resources page for tutorials, guides, etc.

---

# Git Concepts

* Git tracks the history of a folder and its contents
  - the folder is called a *repository* or *repo*
* Versioning happens when you take a snapshot of repo:
  - the snapshot is called a *commit*

---

# Commits

* you *manually* include new files or files that have changed
  - (Git stores information about changes)
* each commit requires a *commit message*
  - ideally summarizing the changes
* commits also record *who* and *when* (automatically)

---

# GitHub

* a cloud platform that hosts Git repos
* free *public* and *private* repos
* size limits:
  - 1 GB for each repo
  - 100 MB for individual files
* [GitHub Education](https://education.github.com/pack) has some benefits
  - but they have recently been making more features available for free anyway

---

class: inverse, center, middle

# GitHub in a Web Browser

---

# Using GitHub

* https://github.com
* create an account / log in
* create a new repo
* upload files
* write a commit message

---

class: center, middle

# DEMO

---

class: inverse, center, middle

# Using Git Clones

---

# Synchronizing Repos

* Git repos can exist in multiple locations:
  - *clone* = "make a copy of this Git repo"
  - *push* = "send new commits from here to somewhere else"
  - *pull* = "retrieve new commits from somewhere else to here"

---

# Working with Git clones

* you need the `Git` software
  - [instructions from Software Carpentry](https://carpentries.github.io/workshop-template/#git)
  - [guide to command-line usage](https://swcarpentry.github.io/git-novice/)  
    (too advanced for this workshop)

--

* [GitHub Desktop](https://desktop.github.com/)
  - graphical interface for Git
  - also syncs your computer with GitHub
  
---

class: center, middle

# DEMO

---

class: inverse, center, middle

# Best Practices for Git Repos

---

# Readme

* a `README.md` file in the top directory
  - what is the project 

---

# License

* a `LICENSE` file in the top directory
  - how other people can *use* your content
  - by default, in the US, you have full rights on creative works
* select based on content type:
  - MIT (permissive, code)
  - CC-BY (permissive, non-code)
  - CC0 (public-domain, data)
  
---

# Adding a common open source license

* when creating a repo on [github.com](https://github.com) or in GitHub Desktop, you can select a license template

* in the web browser, adding a file named `LICENSE` brings up a button to "Choose a license template"

---

# Thanks

* Let me know what content you'd like to see
* Contact me for additional questions or consultation requests!
* Check back in on the libguide for more modules and contact info:
  - https://guides.uflib.ufl.edu/reproducibility
